---
title: "Getting Euroleague Basketball stats using euroleagueR"
author: "Jason Zivkovic"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Euroleague Basketball stats using euroleagueR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message=FALSE,
  warning=FALSE
)
```


## Overview

This vignette will outline how to use the functions in the `euroleagueR` package to allow users to obtain clean and tidy data for matches played in the Euroleague basketball competition.


## Installation

Currently, this library is not yet on CRAN, so you can install the dev version of [**```euroleagueR```**](https://github.com/JaseZiv/euroleagueR/) from [GitHub](https://github.com/JaseZiv/euroleagueR) with:

```{r gh-installation, eval=FALSE}
# install.packages("remotes")
remotes::install_github("JaseZiv/euroleagueR")
```



```{r libraries}
library(euroleagueR)
library(dplyr)
```


## Usage


The following code will show you how to get various different statistics for the Euroleague.


### Match Results

To get results of matches players in the Euroleague tournament since the 2001-2001 season, use the `euroleague_results()` function. 

This data set will also be important to load in together with any of the following data sets to provide match-level metadata (date, venue, scores, phase type, etc)

```{r match_results_wide}
results <- euroleague_results()
dplyr::glimpse(results)
```


#### Joining keys

To be able to join match-level metadata, `season_code` and `code` are the two columns needed for joining on to any of the below listed data sets.


### Team Box Scores

To get box scores at the *team* level, use the `euroleague_team_box()` function. This can return team box score data since the 2010-2011 season, in long form, meaning there will be two rows for each game - one row per team. 

Note, not all fields will contain data for earlier seasons.

```{r team_box}
# the below will get all available match's team box score data
team_box <- euroleague_team_box(season = c(2010:2022))
glimpse(team_box)
```


Now say we want to be able to join on extra information - say the scores, home and away teams, venue, group name, etc, the below code can be used:

```{r rich_team_box}
rich_team_box <- results |> 
  select(code, season_code, group_name, phase_type_name, date, code_home, code_away, score_home, score_away, venue_name, coach_name_home, coach_name_away) |> 
  inner_join(team_box, by = c("season_code", "code"), multiple = "all")
```



### Player Box Scores

To get box scores at the *player* level, use the `euroleague_player_box()` function. This can return player box score data since the 2010-2011 season. Note, not all fields will contain data for earlier seasons.

```{r player_box}
player_box <- euroleague_player_box(season = c(2020:2022))
glimpse(player_box)
```


### Play-by-Play data

To get play-by-play data, use the `euroleague_pbp()` function. This can return play-by-play data since the 2010-2011 season. Note, not all fields will contain data for earlier seasons.

```{r pbp}
pbp <- euroleague_pbp(season = c(2021:2022))
glimpse(pbp)
```



### Shots data

To get shot location and type data, use the `euroleague_shots()` function. This can return shooting data since the 2010-2011 season. Note, not all fields will contain data for earlier seasons.

```{r shots}
shots <- euroleague_shots(season = 2022)
glimpse(shots)
```

